Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2015-06-03 unstructured.tex
2015.06.03 (Mema & Vassilis) Mema's latest pass. Also added paper for citation LCCLS2002.

In \emph{iterative deepening}, 
%(see Figure~\ref{figure:dbfs:iterdeep})
the search is
performed on a BFS tree with multiple preset depths. 
The depth limit is iteratively increased by the source node for each query, 
based on the quality of the results. 
The source node may issue a new request with increased depth limit
that will trigger the nodes at the last visited depth level to resume the
search thus avoiding restart of the entire search process and consequently 
reducing the load on the nodes of the upper levels of the tree. 
Its major drawback is the delay between successive iterations, as the
source-node has to examine the results at each attempt before deciding to
either quit or ``unfreeze'' the query.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \paragraph*{ \bf Delay Aware P2P System}
A \emph{delay--aware} \p\ system termed {\sl DAPS} 
was introduced in~\cite{ZL2005}.
{\sl DAPS}  seeks to attain reduced look-up times by dividing 
peer routing tables into several sectors of increasing delay. 
The source node that issues the query designates the delay 
boundary it may tolerate, providing so a \emph{pruning factor}.
In this context, user requests are forwarded only to 
nodes whose expected delay is less than or equal to the indicated boundary. 
{\sl DAPS} primarily focuses on user ``experience'' and deploys an
end-to-end delay monitoring mechanism that may enable
the clustering of routing tables. 
In a dynamic environment, the formation 
of very accurate such routing tables might be however an elusive goal.
%With the clustered
%routing tables and the loose organization of DAPS' overlay network it is
%considered by is be tween structured and unstructured.
%%
In terms of the three performance criteria, {\sl DAPS} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Does not check for close by nodes, it just sorts the contents of a node's
% routing table.
% Also in a high churn environment a lot of time could be spend probing for
% distances, resorting routing tables and creating sectors. Moreover it just
% prunes distant neighbor
low &
% Sorting routing table entries is done locally so it is not affecting.
low &
%
medium
\end{tabular}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Gia}
The key objective of the {\sl Gia} system~\cite{CRBLS2003} is 
to help alleviate the scalability omnipresent
in unstructured \p\ file-sharing systems.
At first, {\sl Gia} replaced {\sl Gnutella}'s blind flooding 
with random walks.
Although this adoption was 
a step in the right direction \cite{LCCLS2002},
issuing a single copy of the query within the network
effectively reduces the search scope and may negatively affect 
the success rate of the query in question.
To overcome this limitation, {\sl Gia} introduces
a token-based flow control mechanism 
that gradually redirects queries to nodes which are more
likely to answer. This flow control mechanism also helps prevent node
overloading as each peer ``announces'' the number of query requests it 
can handle, in terms of tokens, to its neighbors; to this end, 
peers only forward query requests to nodes that they previously
received tokens from. Further refinement to the search mechanism is the
support for \emph{one--hop replication} of pointers to content.
{\sl Gia} also acknowledges the heterogeneity in peer bandwidth,
processing power, disk speed, etc., of \p\ nodes and uses this
information when connecting nodes to each other.
By using a topology adaptation algorithm, 
{\sl Gia} places low capacity nodes within 
short proximity to peers with high performance features.
This topology adaptation algorithm is based on the metric 
each node maintains about its satisfaction --ranging between $0$ and $1$--
for the neighbors it finds itself associated with. 
Through message exchange, a peer can establish new connections 
or drop superseded ones in order to improve its satisfaction.
Despite the fact that {\sl Gia}'s topology adaptation algorithm 
that primarily focuses on the satisfaction of peers improves the network's
scalability, it falls short in addressing the mismatch problem
as considerations for the underlying network are not handled in explicit
terms \cite{LXLNZ2005}. Moreover it reduces the search scope, exhibits poor
performance in the worst case when matching data is not found quickly
\cite{PR2004,HJ2004} and can potentially build  disconnected topologies
\cite{MBL2006}.
%
In terms of the three stated criteria, {\sl Gia}'s approach is as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Reduces the search scope.
medium &
% Only one copy of the query is issued to the network.
% A token based flow control mechanism redirects queries to nodes that are most
% likely to fulfill them.
low &
% Even thought the low overhead scalability is reduced by the reducing of
% search scope.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
Another problem with topology unaware
systems is the duplication of messages due to cycles in the network graph.
Interestingly, such cycles appear even along the correct forwarding path.
\cite{ZKB2008} focuses on this exact deficiency of overlay networks and
introduces the \emph{Distributed Cycle Minimization Protocol
(DCMP)}, a dynamic, fully distributed method that removes cycles;
this is accomplished without sacrificing
overlay connectivity, resilience and other key properties of unstructured
\p\ architectures and by avoiding a hierarchical organization of peers. 
Once a cycle is detected in {\it DCMP}, the most powerful node in that cycle 
is elected as the \emph{Gate Peer} and the cycle is then broken in that place
that it will result in the minimization of the distance between the
\emph{Gate Peer} and all other nodes that are currently part of the cycle. 
This process is managed by using two specialized message types 
namely, \emph{Information Collection Message (ICM)} and \emph{Cut Message (CM)}. 
\emph{DCMP} bases its operation on messages whose travel is limited by an
imposed \emph{TTL} value (max set to $7$ for most cases). This inherently limits
the protocol in detecting cycles that span for more than \emph{TTL} nodes.
Even though cycle elimination does improve network performance, 
it cannot  directly contribute to the solution of the topology mismatch
problem, while its performance in a high traffic and churn environments and its
ability to retain important characteristics of a {\sl Gnutella}-like topology (i.e.,
like degree distribution, average peer distance, diameter e.t.c.) has been
doubted~\cite{CSG2010}. 
On the other hand, even though some of the overhead
saved by the duplicate message reduction is spent to control the cycle cutting
mechanism, it is reported that it does a lot better in that area
due to its ``lazy'' broadcasting of control messages, as opposed to, for example,
LTM's periodic approach \cite{ZKB2008}.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
%\caption{Cycle elimination methods in DCMP}
%\label{figure:dcmp}
%\end{figure}
%
% TODO: SOME DISCUSSION
%Actually, the duplicate packet problem seems to hurt more the active nodes;
%those with higher capacities and bandwidth that contribute the most to the
%network.
The expected \emph{DCMP} behavior as far as our three criteria is:
%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% It returns more result than LTM according to experiments in
% \cite{ZKB2008}
medium &
% Duplicates are an important fraction of the redundant overhead imposed by
% topology mismatch and thus DCMP minimizes this as possible.
% It is reported though that still stays for example one to two orders of
% magnitude less than LTM due to its ``lazy'' broadcasting of control messages
% (as opposed to LTM's periodic approach) \cite{ZKB2008}.
low &
% fully distributed method
high
\end{tabular}
}
\end{center}

@@

In \cite{CS2002} and \cite{LCCLS2002}, replication is used as a way to improve
inefficient blind search. The observation is that as the number of replicas
increases in the network, it would be relatively easy to locate items even if
search is random. Concretely, three three approaches for replication allocation
are discussed: \emph{uniform}, \emph{proportional}, and
\emph{square-root replication}.

In the uniform model, all objects are replicated without taking into account
the query distribution,
while in the proportional, objects are replicated analogously to their query
rate, so that frequently queried items can be found more often. Experimental
outcomes indicate that the uniform allocation minimizes the maximum search
length and so it can reduce the time spent on unsuccessful searches. The
proportional strategy effectively decreases the search time for popular items,
but suffers when needing to locate the rare entities. When it comes to expected
successful search size, it is the same for for both uniform and proportional
models and any approach in between to always behave much better. Finally, the
square-root model is an attempt to find the golden ratio between uniform and
proportional allocation models. It suggests setting the number of object
replicas to the square root of the searching rate for an object\footnote{An
interesting application of the square-root principle has also been explored
in the context of topology reconstruction, by ensuring node degree to be
proportional to the square-root of its content popularity \cite{C2005}.}.
Theory in \cite{CS2002} suggest that square-root minimizes the expected search size on
successful queries, claim that is supported by simulations \cite{LCCLS2002}.

In \cite{LCCLS2002}, a \emph{k-walker} query algorithm is presented and shown
that using it along side square-root allocation results in performance similar to
{\sl Gnutella}'s query flooding method but incurring up to two
orders of magnitude less network traffic. Also checks three strategies for
replica placement, namely \emph{owner replication} where an object is replicated
at the requesting node, \emph{path replication} that caches the object along the
path from the requesting to the providing node and \emph{random replication}
where the replication is taking place randomly on visited nodes. Observations in
this work are not directly applicable to structured DHTs, because it is assumed
that the lookup time for an object depends only on the number of replicas and
not the placement strategy \cite{RS2004}.

@@

{\sl Narada}~\cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks capable of 
application-layer multicast communications without requiring IP multicast
infrastructure at the network layer.
Although IP--multicast would present a choice in implementing
{\sl Narada}, it is in general considered that it violates the 
stateless design of the current Internet.
Despite the fact the {\sl Narada} was not designed as \p\ system per se,
it was the first (along with \emph{Scattercast} \cite{C2000}) to consider the
feasibility of overlay-based, application-layer services over the Internet while
taking into account bandwidth and latency properties of the underlying physical
infrastructure. In the context of this work, the inefficiency caused by the
topology mismatch is attempted to be tackled by building a highly connected
graph, termed \emph{mesh},
with each source featuring its own minimum spanning tree. A gossip-protocol was
deployed for the creation of these spanning trees~\cite{LYL2008}.
%as shown in Figure~\ref{figure:narada}.
The graph and trees are dynamically updated as nodes keep joining 
or departing the network. The protocol aims to ease the physical link stress,
the overall resource usage as well as the relative delay among end-systems. 
Unfortunately, the main limitation for {\sl Narada} is that 
although it works reasonably well for small groups, 
it does not scale well for larger networks. 
Hence, it is not a suitable choice for potentially large \p\ networks:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% The overhead of Narada is proportional to the multicast group size
high &
% Maintaining the mesh for large multicast groups is expensive.
low
\end{tabular}
}
\end{center}

@@

Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
optimize the usage of the underlying physical resources and operates in
$2$ phases:
\emph{Selective Flooding} and \emph{Active Topology}.
In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
its immediate neighbors so that queries do not flood the entire network and at
the same time preventing their search scope from shrinking. This way some
neighbors become non--flooding. During \emph{Active Topology}, each peer
replaces independently such non--flooding neighbors, with closer nodes as an attempt
to revise the overlay links so that they can reflect more closely the physical
network topology. Picking a replacement out of these non-flooding
neighbors follows a random policy called \emph{Randomized AT} algorithm.
To accomplish the above actions, a peer has to keep track of 
its communication costs with all its neighbors (e.g., network delays)
as well as the costs between any pair of neighbors. The \emph{randomized AT}
algorithm is applied by a source peer
every time its neighbor list is updated or an updated neighbor cost 
table is received.
%%
%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITHM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
%             IS AOTO)
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbors. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbors.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbors has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbor replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.
%%
\emph{AOTO}'s performance regarding the three criteria is as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multicast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Adaptive Connection Establishment (ACE)}
The \emph{Adaptive Connection Establishment (ACE)} 
approach~\cite{LZXN2004} uses the network delay as a metric 
to estimate the costs between nodes. Each peer computes the costs to its
logical one-hop neighbors and forms a \emph{neighbor cost table (NCT)} 
using a special routing message type. 
Any pair of neighboring peers exchange their \emph{NCT}s 
and so a minimal overlay topology can be formed. Based on the obtained
\emph{NCT}s a minimum spanning tree for each peer and its one-hop neighbors is
created. Finally, neighbors located physically far away 
are replaced by physically-closer counterparts. 
In particular, a peer $S$ iteratively probes the distance $d$ between
itself and every of its non-flooding neighbor nodes $G$ as well as
the distance between $S$ and $G$'s neighbors designated as $H$.
If $d_{SG} > d_{SH}$, then link $SG$ is replaced by link $SH$. If, on the other
hand, $d_{SG} < d_{SH}$ then also $d_{GH}$ and then we have the following
options. Either $d_{SH} < d_{GH}$ in which case link $SH$ is kept or
$d_{SH} > d_{GH}$ in which case $S$ will keep probing other $G$'s direct
neighbors.
%
The above optimization is conducted within $1$-neighbor
closure using as base, a peer and checking all its direct neighbors.
Evidently the scope of such an operation could be extended.
Should  a larger scope be used, a better topology matching can be obtained 
at a greater computational overhead.
%
% TODO: SOME DISCUSSION
%
%Simulations in \cite{liu_acesims_2004} show that the average scope of each
% query to cover the same scope of nodes is reduced by about 65 percent without
%losing any autonomy feature, while the average response time can be reduced by
%35 percent. Larger diameter topologies lead to better topology optimization
%rate but also to higher communication and computation overhead. It was also
%found that it is more effective in higher connectivity dense topologies.
%Compared to LTM, it comes short of convergence speed. In
%\cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and
%response time (80 percent) to message queries without shrinking the search
%scope.Last but
%not least, it is concluded that work must be done on incorporating a more
%sophisticated selection policy for candidate non-flooding peers.
%

@@

\begin{enumerate}
  \item If $Q$ is already a direct neighbor of $S$ then we check 
	the distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$. 
	If either $d_{SQ}$ or $d_{SP}$ is the longest, then the longest
  	link will be placed in a \emph{will-cut} list\footnote{A peer will not send or
  	forward queries to connections in its will-cut list but will preserve them
  	for some time in order not to effect ongoing response messages travelling the
  	inverse path.}. 
	If $d_{PQ}$ is the longest, then nothing is done by $S$; being
  	fully distributed, neighbor comparison and selection process, will have the
  	opportunity to handle $PQ$ link when initiated by peers $P$ or $Q$.

@@

In what follows, we outline how the surveyed protocols 
use elements of the above four methodologies. 
We then  offer a summary qualitative comparison for all surveyed
approaches applicable for \emph{unstructured}~\p~\emph{systems}.
Specifically, Table~\ref{unstructured:table} summarizes, for each
effort, the methodologies it incorporates, its special highlights as well as 
a rough estimation of the pros and cons in its implementation. 
Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
and \ref{unstructured:plot:scalability} pictorially compare the
surveyed techniques in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability};
we defined these three criteria in Subsection~\ref{background:motivation}
and we collectively used them as a yardstick to qualitatively 
evaluate the presented approaches.

>>>>

In \emph{iterative deepening}, 
%(see Figure~\ref{figure:dbfs:iterdeep})
the search is
performed on a BFS tree with multiple preset depths. 
The depth limit is iteratively increased by the source node for each query, 
based on the quality of the results. 
The source node may issue a new request with increased depth limit
that will trigger the nodes at the last visited depth level to resume the
search thus avoiding restart of the entire search process and consequently 
reducing the load on the nodes of the upper levels of the tree. 
Its major drawback is the delay between successive iterations, as the
source-node has to examine the results with each attempt before deciding to
either quit or ``unfreeze'' the query.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \paragraph*{ \bf Delay Aware P2P System}
A \emph{delay--aware} \p\ system termed {\sl DAPS} 
was introduced in~\cite{ZL2005}.
{\sl DAPS}  seeks to attain reduced look-up times by dividing 
peer routing tables into several sectors of increasing delay. 
The source node that issues the query designates the delay 
boundary it may tolerate, by providing a \emph{pruning factor}.
In this context, user requests are forwarded only to 
nodes whose expected delay is less than or equal to the indicated boundary. 
{\sl DAPS} primarily focuses on user ``experience'' and deploys an
end-to-end delay monitoring mechanism that may enable
the clustering of routing tables. 
In a dynamic environment, the formation of such 
accurate routing tables may be an elusive goal.
%With the clustered
%routing tables and the loose organization of DAPS' overlay network it is
%considered by is be tween structured and unstructured.
%%
In terms of the three performance criteria, {\sl DAPS} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Does not check for close by nodes, it just sorts the contents of a node's
% routing table.
% Also in a high churn environment a lot of time could be spend probing for
% distances, resorting routing tables and creating sectors. Moreover it just
% prunes distant neighbor
low &
% Sorting routing table entries is done locally so it is not affecting.
low &
%
medium
\end{tabular}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Gia}
The key objective of the {\sl Gia} system~\cite{CRBLS2003} is 
to help alleviate the scalability omnipresent
in unstructured \p\ file-sharing systems.
At first, {\sl Gia} replaced {\sl Gnutella}'s blind flooding 
with random walks.
Although this adoption was 
a step in the right direction \cite{LCCLS2002},
issuing a single copy of the query within the network
effectively reduces the search scope and may negatively affect 
the success rate of the query in question.
To overcome this limitation, {\sl Gia} introduces
a token-based flow control mechanism 
that gradually redirects queries to nodes which are more
likely to answer. This flow control mechanism also helps prevent node
overloading as each peer ``announces'' the number of query requests it 
can handle, in terms of tokens, to its neighbors; to this end, 
peers only forward query requests to nodes that they previously
received tokens from. Further refinement to the search mechanism is the
support for \emph{one--hop replication} of pointers to content.
{\sl Gia} also acknowledges the heterogeneity in peer bandwidth,
processing power, disk speed, etc., of \p\ nodes and uses this
information when connecting nodes to each other.
By using a topology adaptation algorithm, 
{\sl Gia} places low capacity nodes within 
short proximity to peers with high performance features.
This topology adaptation algorithm is based on the metric 
each node maintains about its satisfaction --ranging between $0$ and $1$--
for the neighbors with which it finds itself associated. 
Through message exchanges, a peer can establish new connections 
or drop superseded ones to improve its satisfaction.
Despite the fact that {\sl Gia}'s topology adaptation algorithm 
improves the network's
scalability, it falls short in addressing the mismatch problem
as considerations for the underlying network are not handled in explicit
terms \cite{LXLNZ2005}. Moreover it reduces the search scope, exhibits poor
performance in the worst case when matching data is not found quickly
\cite{PR2004,HJ2004} and can potentially build  disconnected topologies
\cite{MBL2006}.
%
In terms of the three stated criteria, {\sl Gia}'s approach is as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Reduces the search scope.
medium &
% Only one copy of the query is issued to the network.
% A token based flow control mechanism redirects queries to nodes that are most
% likely to fulfill them.
low &
% Even thought the low overhead scalability is reduced by the reducing of
% search scope.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
Another problem with topology unaware
systems is the duplication of messages due to cycles in the network graph.
Interestingly, such cycles appear even along the correct forwarding path.
\cite{ZKB2008} focuses on this exact deficiency of overlay networks and
introduces the \emph{Distributed Cycle Minimization Protocol
(DCMP)}, a dynamic, fully distributed method that removes cycles;
this is accomplished without sacrificing
overlay connectivity, resilience and other key properties of unstructured
\p\ architectures and by avoiding a hierarchical organization of peers. 
Once a cycle is detected in {\it DCMP}, the most powerful node in that cycle 
is elected as the \emph{Gate Peer} and the cycle is then broken in that place
that it will result in the minimization of the distance between the
\emph{Gate Peer} and all other nodes that are currently part of the cycle. 
This process is managed by using two specialized message types 
namely, \emph{Information Collection Message (ICM)} and \emph{Cut Message (CM)}. 
\emph{DCMP} bases its operation on messages whose travel is limited by an
imposed \emph{TTL} value (max set to $7$ for most cases). This inherently limits
the protocol in detecting cycles that span for more than \emph{TTL} nodes.
Even though cycle elimination does improve network performance, 
it cannot  directly contribute to the solution of the topology mismatch
problem, while its performance in high-traffic and churn environments and its
ability to retain important characteristics of a {\sl Gnutella}-like topology (i.e.,
degree distribution, average peer distance, diameter e.t.c.) has been
doubted~\cite{CSG2010}. 
On the other hand, overhead is greatly reduced 
due to its ``lazy'' broadcasting of control messages, as opposed to, for example,
LTM's periodic approach \cite{ZKB2008}.
%
%\begin{figure}
%\centering
%  \includegraphics[scale=0.4]{img/dcmp.jpeg}
%\caption{Cycle elimination methods in DCMP}
%\label{figure:dcmp}
%\end{figure}
%
% TODO: SOME DISCUSSION
%Actually, the duplicate packet problem seems to hurt more the active nodes;
%those with higher capacities and bandwidth that contribute the most to the
%network.
The expected \emph{DCMP} behavior as far as our three criteria is:
%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% It returns more result than LTM according to experiments in
% \cite{ZKB2008}
medium &
% Duplicates are an important fraction of the redundant overhead imposed by
% topology mismatch and thus DCMP minimizes this as possible.
% It is reported though that still stays for example one to two orders of
% magnitude less than LTM due to its ``lazy'' broadcasting of control messages
% (as opposed to LTM's periodic approach) \cite{ZKB2008}.
low &
% fully distributed method
high
\end{tabular}
}
\end{center}

@@

In \cite{CS2002} and \cite{LCCLS2002}, replication is used as a way to improve
inefficient blind search. The observation is that as the number of replicas
increases in the network, it would be relatively easy to locate items even if
search is random. Concretely, three three approaches for replication allocation
are proposed: \emph{uniform}, \emph{proportional}, and
\emph{square-root replication}.

In the uniform model, all objects are replicated without taking into account
the query distribution,
while in the proportional, objects are replicated analogously to their query
rate, so that frequently queried items can be found more often. Experimental
outcomes indicate that the uniform allocation minimizes the maximum search
length and so it can reduce the time spent on unsuccessful searches. The
proportional strategy effectively decreases the search time for popular items,
but suffers when needing to locate rare items. For allocation ratio ``between''
1 and the ratio of the query rates, the authors
prove that the average performance on successful queries is always better
than in uniform and proportional models. Finally, with the
square-root model they attempt to find the golden ratio in ``between'', by
suggesting to set the number of object
replicas to the square root of the searching rate for an object\footnote{An
interesting application of the square-root principle has also been explored
in the context of topology reconstruction, by ensuring node degree to be
proportional to the square-root of its content popularity \cite{C2005}.}.
Theory in \cite{CS2002} suggest that square-root minimizes the expected search size on
successful queries, a claim that is supported by simulations \cite{LCCLS2002}.

Lv et al.~\cite{LCCLS2002}, present a \emph{k-walker} query algorithm and show
that using it along side square-root allocation results in performance similar to
{\sl Gnutella}'s query flooding method but incurs up to two
orders of magnitude less network traffic. Also checks three strategies for
replica placement, namely \emph{owner replication} where an object is replicated
at the requesting node, \emph{path replication} that caches the object along the
path from the requesting to the providing node and \emph{random replication}
where the replication is taking place randomly on visited nodes. Observations in
this work are not directly applicable to structured DHTs, because it is assumed
that the lookup time for an object depends only on the number of replicas and
not the placement strategy \cite{RS2004}.

@@

{\sl Narada}~\cite{CRZ2000,CRSZ2001,CRSZ2002} is a generic
protocol for creating self-adapting overlay networks capable of 
application-layer multicast communications without requiring IP multicast
infrastructure at the network layer.
Although IP--multicast would present a choice in implementing
{\sl Narada}, it is in general considered that it violates the 
stateless design of the current Internet.
Despite the fact the {\sl Narada} was not designed as \p\ system per se,
it was the first (along with \emph{Scattercast} \cite{C2000}) to consider the
feasibility of overlay-based, application-layer services over the Internet while
taking into account bandwidth and latency properties of the underlying physical
infrastructure. To overcome the inefficiency caused by 
topology mismatch, Narada builds a highly connected
graph, termed \emph{mesh},
with each source featuring its own minimum spanning tree. A gossip-protocol was
deployed for the creation of these spanning trees~\cite{LYL2008}.
%as shown in Figure~\ref{figure:narada}.
The graph and trees are dynamically updated as nodes keep joining 
or departing the network. The protocol aims to ease the physical link stress,
the overall resource usage as well as the relative delay among end-systems. 
Unfortunately, the main limitation for {\sl Narada} is that 
although it works reasonably well for small groups, 
it does not scale well for larger networks. 
Hence, it is not a suitable choice for potentially large \p\ networks:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% The overhead of Narada is proportional to the multicast group size
high &
% Maintaining the mesh for large multicast groups is expensive.
low
\end{tabular}
}
\end{center}

@@

Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
optimize the usage of the underlying physical resources and operates in
$2$ phases:
\emph{Selective Flooding} and \emph{Active Topology}.
In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
its immediate neighbors so that queries do not flood the entire network 
while preventing their search scope from shrinking. This way some
neighbors become non--flooding. During \emph{Active Topology}, each peer
replaces independently such non--flooding neighbors, with closer nodes as an attempt
to revise the overlay links so that they can reflect more closely the physical
network topology. Picking a replacement out of these non-flooding
neighbors follows a random policy called \emph{Randomized AT} algorithm.
To accomplish the above actions, a peer has to keep track of 
its communication costs with all its neighbors (e.g., network delays)
as well as the costs between any pair of neighbors. The \emph{randomized AT}
algorithm is applied by a source peer
every time its neighbor list is updated or an updated neighbor cost 
table is received.
%%
%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITHM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
%             IS AOTO)
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbors. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbors.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbors has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbor replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.
%%
\emph{AOTO}'s performance regarding the three criteria is as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multicast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Adaptive Connection Establishment (ACE)}
The \emph{Adaptive Connection Establishment (ACE)} 
approach~\cite{LZXN2004} uses the network delay as a metric 
to estimate the costs between nodes. Each peer computes the costs to its
logical one-hop neighbors and forms a \emph{neighbor cost table (NCT)} 
using a special routing message type. 
Any pair of neighboring peers exchange their \emph{NCT}s 
and so a minimal overlay topology can be formed. Based on the obtained
\emph{NCT}s a minimum spanning tree for each peer and its one-hop neighbors is
created. Finally, neighbors located physically far away 
are replaced by physically-closer counterparts. 
In particular, a peer $S$ iteratively probes the distance $d$ between
itself and each of its non-flooding neighbor nodes $G$ as well as
the distance between $S$ and $G$'s neighbors designated as $H$.
If $d_{SG} > d_{SH}$, then link $SG$ is replaced by link $SH$. If, on the other
hand, $d_{SG} < d_{SH}$ then we have the following
options. Either $d_{SH} < d_{GH}$ in which case link $SH$ is kept or
$d_{SH} > d_{GH}$ in which case $S$ will keep probing other $G$'s direct
neighbors.
%
The above optimization is conducted within $1$-neighbor
closure using as base, a peer and checking all its direct neighbors.
Evidently the scope of such an operation could be extended.
Should  a larger scope be used, a better topology matching can be obtained 
at a greater computational overhead.
%
% TODO: SOME DISCUSSION
%
%Simulations in \cite{liu_acesims_2004} show that the average scope of each
% query to cover the same scope of nodes is reduced by about 65 percent without
%losing any autonomy feature, while the average response time can be reduced by
%35 percent. Larger diameter topologies lead to better topology optimization
%rate but also to higher communication and computation overhead. It was also
%found that it is more effective in higher connectivity dense topologies.
%Compared to LTM, it comes short of convergence speed. In
%\cite{ni_mismatch_2004} shows reduction of both total traffic (90 percent) and
%response time (80 percent) to message queries without shrinking the search
%scope.Last but
%not least, it is concluded that work must be done on incorporating a more
%sophisticated selection policy for candidate non-flooding peers.
%

@@

\begin{enumerate}
  \item If $Q$ is already a direct neighbor of $S$ then we check 
	the distances $d_{SQ}$, $d_{SP}$ and $d_{PQ}$. 
	If either $d_{SQ}$ or $d_{SP}$ is the longest, then the longest
  	link will be placed in a \emph{will-cut} list\footnote{A peer will not send or
  	forward queries to connections in its will-cut list but will preserve them
  	for some time in order not to affect ongoing response messages traveling the
  	inverse path.}. 
	If $d_{PQ}$ is the longest, then nothing is done by $S$; being
  	fully distributed, neighbor comparison and selection process, will have the
  	opportunity to handle $PQ$ link when initiated by peers $P$ or $Q$.

@@

In what follows, we outline how the surveyed protocols 
use elements of the above four methodologies. 
We then  offer a summary qualitative comparison for all surveyed
approaches applicable for \emph{unstructured}~\p~\emph{systems}.
Specifically, Table~\ref{unstructured:table} summarizes, for each
effort, the methodologies it incorporates, its special highlights as well as 
a brief presentation of the pros and cons in its implementation. 
Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
and \ref{unstructured:plot:scalability} pictorially compare the
surveyed techniques in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability};
we defined these three criteria in Subsection~\ref{background:motivation}
and we collectively used them as a yardstick to qualitatively 
evaluate the presented approaches.
